---
title: "DE CARLO"
author: "Mégane"
date: "2023-04-20"
output: html_document
---

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
library("readr")
library("viridis")
library("scales")
library("lubridate")
library("gsignal")
library("FactoMineR")
library("factoextra")
library("gplots")
library("heatwaveR")
library("dplyr")
library("imputeTS")
library("Metrics")
library("tidyr")
library("tseries")
library("forecast")
library("tsDyn")
library("tsoutliers")
library("seacarb")
library("lmodel2")
library("lmtest")
library("ggExtra")
library("corrplot")
library("ggcorrplot")
library("marelac")
library("kableExtra")
library("stringr")
library("berryFunctions")
```


#### *Data wind speed 10m (m/s)*:  
  
→ Azur buoy (model) - 2009-2022  
→ EOL buoy (model) - 2011-2022  
  
Frequency : Hourly  
Unity : m/s

```{r Importation data wind speed, echo=FALSE, warning=FALSE, message=FALSE}

#importation data wind speed (model) of Azur buoy (2009-2022) :
Wind_Azur <- read_csv("../Data/Wind_speed_10m_Azur_buoy/model_wind_speed_AZUR.csv")

#visualization NA 
#ggplot_na_distribution(Wind_Azur$speed_ms, x_axis_labels=Wind_Azur$Date, xlab="",
#                       ylab="Wind speed (m/s)", title="Wind speed 10m (m/s) of Azur buoy (2009-2022)")

# pas de NA, complete series

##

#importation data wind speed (model) of Azur buoy (fin 2011-2022) :
Wind_EOL <- read_csv("../Data/model_wind_speed_EOL.csv")

#visualization NA 
# ggplot_na_distribution(Wind_EOL$speed_ms, x_axis_labels=Wind_EOL$Date, xlab="",
#                        ylab="Wind speed (m/s)", title="Wind speed 10m (m/s) of EOL buoy (2011-2022)")

# pas de NA, complete series

##
```



### **Histogram of wind speed distribution**  

```{r Histogram of wind speed distribution, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=4, fig.align='center', fig.show='hold'}

#plot 1
Wind_Azur %>% 
  ggplot(aes(x=speed_ms)) +
  geom_histogram(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
  ggtitle("Wind speed distribution (Azur buoy)") + 
  scale_y_continuous(name="Distribution") +
  scale_x_continuous(name="", breaks = c(0, 5, 10, 15, 20))
##

#plot 2
Wind_EOL %>% 
  ggplot(aes(x=speed_ms)) +
  geom_histogram(fill="#ED7F10", color="#e9ecef", alpha=0.8) +
  ggtitle("Wind speed distribution (EOL buoy)") + 
  scale_y_continuous(name="Distribution") +
  scale_x_continuous(name="", breaks = c(0, 5, 10, 15, 20))



```

```{r filter 9 AM, echo=FALSE, warning=FALSE, message=FALSE}

Wind_EOL_9AM <- Wind_EOL %>% 
  mutate(Hour = format(date, format="%H:%M:%S")) %>% 
  dplyr::filter(Hour == "09:00:00") %>% 
  mutate(datetime = format(date, format="%Y-%m-%d"),
         datetime = as.POSIXct(datetime)) %>% 
  select(datetime, speed_ms)

```




#### *Data pCO2 water SOMLIT (µatm)*:  
  
period : jan.2007 - jan.2023 
  
Frequency : Weekly (every tuesday at 09 AM)  
Unity : µatm  
  
  
→ Linear interpolation pCO2 water :  
  
```{r importation data pCO2 water, echo=FALSE, warning=FALSE, message=FALSE}

#period : jan. 2007 - jan. 2023
#measurements every tuesday at 09h AM

SOMLIT_pcO2 <- read_delim("../Data/DATA_SOMLIT_07-23_CLEAN.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)


#surface data :
SOMLIT_pcO2 <- SOMLIT_pcO2 %>% 
  dplyr::filter(depth==0) %>% 
  rename(pCO2_w = pCO2, datetime = sampling_date) %>%
  mutate(salinity = case_when(salinity < 35 ~ NA_real_ , TRUE ~ salinity))



#linear interpolation pCO2 atmos :
impute_pCO2_w <- na_interpolation(SOMLIT_pcO2$pCO2_w, option="linear")

ggplot_na_imputations(SOMLIT_pcO2$pCO2_w, impute_pCO2_w, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "", ylab="pCO2 (µatm)", 
                      title="Water pCO2 (µatm) time series interpolation - period 2007-2022")


#merge wind speed & SOMLIT :
merge1 <- left_join(SOMLIT_pcO2, Wind_EOL_9AM, by="datetime")


```

  
→ Linear interpolation temperature & salinity :  
  

```{r interpolation temp & salinity, echo=FALSE, warning=FALSE, message=FALSE}

impute_temp <- na_interpolation(SOMLIT_pcO2$temperature, option="linear")
impute_sal <- na_interpolation(SOMLIT_pcO2$salinity, option="linear")

ggplot_na_imputations(SOMLIT_pcO2$temperature, impute_temp, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "", ylab="Temp. (°C)", title="Point B - Temperature time series (°C) - period 2007-2022")
ggplot_na_imputations(SOMLIT_pcO2$salinity, impute_sal, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "", ylab="Salinity", title="Point B - Salinity time series - period 2007-2022")


merge1 <- merge1 %>% 
  mutate(impute_temp = impute_temp, impute_sal = impute_sal)


```

```{r interpolation AT & CT, echo=FALSE, warning=FALSE, message=FALSE}

impute_ta <- na_interpolation(SOMLIT_pcO2$ta, option="linear")
#faire pour CT
#impute_CT <- na_interpolation()

ggplot_na_imputations(SOMLIT_pcO2$ta, impute_ta, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "", ylab="AT (µmol/kg)", 
                      title="Point B - Total alkalinity time series (µmol/kg) - period 2007-2022")
#faire pour CT
# ggplot_na_imputations(SOMLIT_pcO2$, impute_CT, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "", ylab="CT", 
#                       title="Point B - Total Carbon time series (µmol/kg) - period 2007-2022")


# merge1 <- merge1 %>% 
#   mutate(impute_ta = impute_ta)

```


#### *Data pCO2 air (µatm)*:  
  
period : jan.2007 - jan.2023 
  
Frequency : Hourly
Unity : µatm  
  
→ Corse Ersa (40m) - 2013-2020  
→ Plateau Rosa (10m) - 2007-2022  
→ Lampedusa (10m) - 2006-2022  
  
```{r importation data pCO2 air, echo=FALSE, warning=FALSE, message=FALSE}

#CORSE ERSA
pCO2_CORSE <- read_delim("../Data/atmospheric_co2_CORSE_RAW_2013_2020.CO2", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

# delate last row :
pCO2_CORSE <- pCO2_CORSE[-65900,]
#46 days of measurements are missing : 2789 - 2743 (theo - obs)

#selection of more or less 09:00 AM data :
pCO2_CORSE <- pCO2_CORSE %>% 
  unite(col='datetime', c('Year', 'Month', 'Day'), sep='-') %>% 
  unite(col='time', c('Hour', 'Minute'), sep=':') %>% 
  dplyr::filter(time == "09:00")

#data sorting :
pCO2_CORSE <- pCO2_CORSE %>% 
  mutate(datetime = format(datetime, format="%Y-%m-%d")) %>% 
  mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%d")) %>%
  rename(CO2_CORSE = co2) %>% 
  mutate(CO2_CORSE = case_when(CO2_CORSE == -999.990 ~ NA_real_ , TRUE ~ CO2_CORSE)) %>% 
  select(datetime, CO2_CORSE)

# 1534 NA
###########################

### PLATEAU ROSA
pCO2_PLATEAU_ROSA <- read_delim("../Data/atmospheric_co2_PLATEAU_ROSA_2007_2022.CO2", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

#selection of 09:00 AM data :
pCO2_PLATEAU_ROSA <- pCO2_PLATEAU_ROSA %>% 
  unite(col='datetime', c('Year', 'Month', 'Day'), sep='-') %>% 
  unite(col='time', c('Hour', 'Minute'), sep=':') %>% 
  dplyr::filter(time == "09:00")


#data sorting :
pCO2_PLATEAU_ROSA <- pCO2_PLATEAU_ROSA %>% 
  mutate(datetime = format(datetime, format="%Y-%m-%d")) %>% 
  mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%d")) %>%
  rename(CO2_PR = co2) %>% 
  mutate(CO2_PR = case_when(CO2_PR == -999.990 ~ NA_real_ , TRUE ~ CO2_PR)) %>% 
  select(datetime, CO2_PR)

# 1262 NA
###########################

#LAMPEDUSA
pCO2_LAMPEDUSA <- read_delim("../Data/atmospheric_co2_LAMPEDUSA_2006_2022.CO2", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

#selection of 09:00 AM data :
pCO2_LAMPEDUSA <- pCO2_LAMPEDUSA %>% 
  unite(col='datetime', c('Year', 'Month', 'Day'), sep='-') %>% 
  unite(col='time', c('Hour', 'Minute'), sep=':') %>% 
  dplyr::filter(time == "09:00")

#data sorting :
pCO2_LAMPEDUSA <- pCO2_LAMPEDUSA %>% 
  mutate(datetime = format(datetime, format="%Y-%m-%d")) %>% 
  mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%d")) %>%
  rename(CO2_LAMP = co2) %>% 
  mutate(CO2_LAMP = case_when(CO2_LAMP == -999.990 ~ NA_real_ , TRUE ~ CO2_LAMP)) %>% 
  select(datetime, CO2_LAMP)

#merge pCO2 air CORSE & Plateau Rosa avec pCO2 water SOMLIT :
merge2 <- left_join(merge1, pCO2_CORSE, by="datetime") #514 NA

merge3 <- left_join(merge2, pCO2_PLATEAU_ROSA, by="datetime") #231 NA

merge3 <- left_join(merge3, pCO2_LAMPEDUSA, by="datetime") #281 NA


```

###### Interpolation of atmospheric pCO2 :  
  
→ Cross interpolation between 3 atmospheric sites : Corsica x Lampedusa x Plateau Rosa  
  
→ Then, linear interpolation :  
*(+ conversion dry mol fraction into µatm)*  
  
```{r conversion dry mol fraction + interpolation data pCO2 air, echo=FALSE, warning=FALSE, message=FALSE}

#pCO2_atmos = interpolation between 3 sites (Corsica x Lampedusa x Plateau Rosa)

merge3 <- merge3 %>% 
  mutate(CO2_CORSE = str_replace_na(CO2_CORSE, replacement = "NA"),
         CO2_PR = str_replace_na(CO2_PR, replacement = "NA"),
         CO2_LAMP = str_replace_na(CO2_LAMP, replacement = "NA"),
         pCO2_atmos_dry = case_when(CO2_CORSE == "NA" ~ CO2_LAMP, TRUE ~ CO2_CORSE),
         pCO2_atmos_dry = case_when(pCO2_atmos_dry == "NA" ~ CO2_PR, TRUE ~ pCO2_atmos_dry))

merge3 <- merge3 %>% 
  mutate(CO2_CORSE = as.numeric(CO2_CORSE),
         CO2_PR = as.numeric(CO2_PR),
         CO2_LAMP = as.numeric(CO2_LAMP),
         pCO2_atmos_dry = as.numeric(pCO2_atmos_dry)) #102 NAs



#conversion dry mol fraction pCO2 air into µatm :
merge3 <- merge3 %>% 
  mutate(pCO2_atmos = x2pCO2(S=35, T=25, Patm=1.0, xCO2=pCO2_atmos_dry))



#linear interpolation pCO2 atmos :
impute_pCO2_atmos <- na_interpolation(merge3$pCO2_atmos, option="linear")

ggplot_na_imputations(merge3$pCO2_atmos, impute_pCO2_atmos, x_axis_labels = merge3$datetime, xlab = "", ylab="pCO2 (µatm)", 
                      title="Atmospheric pCO2 (µatm) time series interpolation - period 2007-2022")

merge3 <- merge3 %>% 
  mutate(impute_pCO2_atmos = impute_pCO2_atmos)

```


```{r calcul delta pCO2, echo=FALSE, warning=FALSE, message=FALSE}

#calculation delta pCO2 = pCO2 water - pCO2 air :
merge3 <- merge3 %>%
  mutate(delta_pCO2 = impute_pCO2_w - impute_pCO2_atmos)


```

  

## Air-surface CO~2~ flux  
  
  
### Flux calculation  
  
→ **negative values = flux into the ocean**  
  
  
Method used : Wanninkhof (2014)  
  

```{r calcul flux, echo=FALSE, warning=FALSE, message=FALSE}

#Calcul of F
tab_flux <- merge3 %>% 
    mutate(Flux_CO2_2014 = 0.00077 * speed_ms^2 * delta_pCO2)


#plot Flux Wanninkhof 2014 :
tab_flux %>% 
  ggplot() + 
  aes(x=datetime, y=Flux_CO2_2014) + 
  geom_line(col="#967C5C") +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Flux (mol of C/m²/y)") +
  ggtitle("CO2 fluxes ocean-atmosphere (Wanninkhof 2014)")

```
Method used : Wanninkhof (1992)  
  
   
Formulas :  
  
-   K600 = 0.266 \* U²  
  
-   SC = 2073.1 - (125.62 \* T) + (3.6276 \* T²) - (0.043219 \* T\^3)  
  
-   k = K600 \* (SC/600)\^0.5  
  
-   K0 = Henry's constant (seacarb)  
  
  
Final formula : F = k \* K0 \* ΔpCO2  
  
unity : mol of C / m² / year  
  
```{r calcul Flux Wanninkhof (1992), echo=FALSE, warning=FALSE, message=FALSE}

tab_flux <- tab_flux %>% 
  mutate(SC = 2073.1 - (125.62 * impute_temp) + (3.6276 * impute_temp^2) - (0.043219 * impute_temp^3),
         K600 = 0.266 * speed_ms^2,
         k = K600 * (SC/600)^0.5,
         rho = rho(S=impute_sal, T=impute_temp, P=0)/1000,
         K0 = (K0(S=impute_sal, T=impute_temp, P=0, Patm=1, warn="n")),
         K0 = (K0 * rho)/1000000, 
         Flux_wan_92 = (k * K0 * delta_pCO2)*10*24*365)

##unity steps :

#K0 en mol/kg/atm
#rho (density of water) en kg/m3
#donc rho/1000 = kg/L
#donc K0*(rho/1000) = mol/l/atm

#K0/10^6 pour passer de mol/L/atm a mol/L/µatm

#Flux*10 pour passer de 10 mol/m²/h a mol/m²/h
#Flux*24*365 pour passer de mol/m²/h a mol/m²/y


#plot Flux wanninkhof 1992 :
tab_flux %>% 
  ggplot() + 
  aes(x=datetime, y=Flux_wan_92) + 
  geom_line(col="#677179") +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Flux (mol/m²/y)") +
  ggtitle("CO2 fluxes ocean-atmosphere (Wanninkhof 1992)")


```
  
Method used : David T. Ho (2023)  
  
Formulas :  
  
-   K600 = 0.143 \* U²  
  
-   SC = 2073.1 - (125.62 \* T) + (3.6276 \* T²) - (0.043219 \* T\^3)  
    or gas_schmidt() function of marelac package (same result)
  
-   k = K600 \* (SC/600)\^0.5  
  
-   K0 = Henry's constant (seacarb)
Final formula : F = k \* K0 \* ΔpCO2  
  
unity : mol of C / m² / year


```{r calcul Flux David T. Ho (2023), echo=FALSE, warning=FALSE, message=FALSE}

tab_flux <- tab_flux %>% 
  mutate(K600_M3 = 0.143 * speed_ms^2,
         Flux_HO = ((K600_M3*(SC/600)^0.5) * K0 * delta_pCO2)*10*24*365)

#plot Flux David T. Ho (2023) :
tab_flux %>% 
  ggplot() + 
  aes(x=datetime, y=Flux_HO) + 
  geom_line(col="#24445C") +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Flux (mol/m²/y)") +
  ggtitle("CO2 fluxes ocean-atmosphere (David T. Ho 2023)")

```
  
Summary table of 3 calculation methods results (Wanninkhof 1992 & 2014, David T. Ho 2023) :  
  
  
```{r 3 methods comparison table, echo=FALSE, warning=FALSE, message=FALSE}

summary_table <- data.frame(tab_flux$datetime, tab_flux$Flux_wan_92, tab_flux$Flux_CO2_2014, tab_flux$Flux_HO)

colnames(summary_table) <- c("Datetime", "Flux_wan_92", "Flux_wan_2014", "Flux_Ho_2023")

summary_table[104:837,]


```

Plot comparing the 3 CO~2~ flux calculation results :  
  

```{r 3 methods comparison - plot, echo=FALSE, warning=FALSE, message=FALSE}

#creation column for legend
tab_flux_long <- tab_flux %>% 
  rename(`Flux W (1992)` = Flux_wan_92, `Flux W (2014)` = Flux_CO2_2014,  `Flux Ho (2023)` = Flux_HO) %>% 
  pivot_longer(cols=c(`Flux W (1992)`, `Flux W (2014)`, `Flux Ho (2023)`), names_to = "Legend", values_to = "value")

#plot
tab_flux_long %>% 
  ggplot() +
  ggtitle("CO2 fluxes ocean-atmosphere (3 methods comparison)") +
  geom_line(aes(x=datetime, y=value, color= Legend), size=0.5) +
  scale_y_continuous(name="CO2 flux (mol/m²/y)") + 
  scale_x_datetime(name="")


```


