---
title: "HF EOL"
author: "Mégane"
date: "2023-03-28"
output: html_document
---

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}

library("readr")
library("viridis")
library("scales")
library("lubridate")
library("gsignal")
library("FactoMineR")
library("factoextra")
library("gplots")
library("heatwaveR")
library("dplyr")
library("imputeTS")
library("Metrics")
library("tidyr")
library("tseries")
library("forecast")
library("tsDyn")
library("tsoutliers")
library("seacarb")
library("lmodel2")
library("lmtest")
library("ggExtra")
library("corrplot")
library("ggcorrplot")
library("marelac")

```

```{r theme for plots, echo=FALSE, warning=FALSE, message=FALSE}

#theme for plots :
Mytheme <- function(size_labs = 6, face_font="plain", ...) {
  theme_bw() +
    theme(axis.text.x = element_text(face=face_font, size=size_labs, color="black"),
          axis.title.x = element_text(face=face_font, size=size_labs, margin=margin(0,0,0,0,"pt")),
          axis.text.y = element_text(face=face_font, color="black", size=size_labs),
          axis.title.y = element_text(face=face_font, size=size_labs),
          axis.ticks.x = element_line(size=0.1),
          axis.ticks.y = element_line(size=0.1),
          axis.ticks.length = unit(1.1, "mm"),
          panel.grid.major = element_line(size = 0.25, color="black", linetype="dotted"),
          aspect.ratio = 1 / 3,
          plot.margin = margin(t = 0, r = 1, b = 0, l = 0, unit = "lines")
    )
}
```



# Data EOL surface : temperature (°C), salinity (psu) and oxygen (mll)

## Part 1 : Trend analysis by residuals

```{r importation Data EOL, echo=FALSE, warning=FALSE, message=FALSE}

#period : septembre 2013-octobre 2022
EOL_RAW <- read_csv("../Data/SEANOE_EOL.csv")  

#tri
EOL_RAW <- EOL_RAW %>% 
  rename(temperature = temp_eol_c, salinity = sal_eol_psu, oxygen = oxy_eol_mll) %>% 
  select(-...1)

#summary(EOL_RAW) 
#temperature : 0 NA
#salinity : 34 NA
# oxygen : 4793

```


##### Interpolation - EOL data

```{r partie interpolation temperature, echo=FALSE, warning=FALSE, message=FALSE}

#visualisation NA

#statsNA(EOL_RAW$temperature) # percentage of NA : 0%

```


```{r partie interpolation salinity, echo=FALSE, warning=FALSE, message=FALSE}

#visualisation NA

#ggplot_na_distribution(EOL_RAW$salinity)
#statsNA(EOL_RAW$salinity) # percentage of NA : 0.07%


#interpolation
#salinity_interpol <- na_interpolation(EOL_RAW$salinity)
#ggplot_na_imputations(x_with_na = EOL_RAW$salinity, x_with_imputations = salinity_interpol)
##

#EOL_RAW$salinity <- na_interpolation(EOL_RAW$salinity)

```


```{r period 2014-04-23 _ 2022-04-23, echo=FALSE, warning=FALSE, message=FALSE}

##Essai periode complete 2014-04-23 _ 2022-04-23 : data toutes les h

#EOL_1422 <- EOL_RAW[c(97:42896),] 
  
##plot

#EOL_1422 %>% 
  #ggplot() +
  #geom_line(aes(x= as.Date(yday(datetime), "1970-01-01"), y=temperature, 
               # group = factor(year(datetime)), 
               # color = factor(year(datetime))), size = 0.75) +
 # scale_colour_viridis_d() +
 # scale_x_date(date_breaks="months", date_labels="%b", name = "") +
 # labs(x="Month",colour="") +
 # theme_bw() +
 # scale_y_continuous(name = "Temperature (°C)")


```


##### [Table of monthly means (2013-2022) - EOL temperature & salinity]{style="color:blue"}

```{r calcul monthly means - temperature & salinity, echo=TRUE, warning=FALSE, message=FALSE}

monthly_means <- ungroup(EOL_RAW) %>% 
  mutate(Month = format(datetime, format="%m")) %>% 
  group_by(Month) %>%
  summarise(
    Salinity_month = mean(salinity, na.rm = TRUE),
    Temperature_month = mean(temperature, na.rm = TRUE))

monthly_means

```

##### [Table of time series anomalies regression analyses (2013-2022) - EOL temperature & salinity]{style="color:green"}

```{r regression analysis - temperature & salinity, echo=TRUE, warning=FALSE, message=FALSE}

EOL_RAW <- EOL_RAW %>% 
  mutate(Month = format(datetime, format="%m"))

anomalies <- left_join(ungroup(EOL_RAW), monthly_means, by = "Month") %>%
  mutate(Salinity_ano = salinity - Salinity_month,
         Temperature_ano = temperature - Temperature_month)

# Regressions and tables of key parameters
var_list <-
  c("salinity",
    "temperature")

lm.test <- vector("list", length(var_list))
 
 for(i in seq_along(var_list)){
     lm.test[[i]] <- lm(reformulate(var_list[i], "datetime"), data = anomalies)
 }


lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = anomalies))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_surf <- reg
slope_temp <- reg_ano_surf[2, 1]

# Regression and table anomalies
var_list <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = anomalies))
})
reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")

row.names(reg) <- var_list
reg_anomalies <- reg

reg_anomalies
```

#### Plot of anomalies : Temperature (°C) (2013-2022) - EOL

```{r plot anomalies + trends - temperature, echo=FALSE, warning=FALSE, message=FALSE}


plot_ano_temp <- ggplot(data = anomalies, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="4 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="", y="Temp. (°C)") +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(anomalies$Temperature_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank()) + 
  ggtitle("EOL - Temperature anomalies + trend")

plot_temperature <- EOL_RAW %>% 
  ggplot() +
  aes(x=datetime, y=temperature) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("EOL time series observations of temperature (°C) 2013-2022")

plot_temperature
plot_ano_temp
```


#### Plot of anomalies : Salinity (2013-2022) - EOL

```{r plot anomalies + trends - salinity, echo=FALSE, warning=FALSE, message=FALSE}


plot_ano_sal <- ggplot(data = anomalies, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="4 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="", y="Temp. (°C)") +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(anomalies$Salinity_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank()) + 
  ggtitle("EOL - Salinity anomalies + trend")

plot_salinity <- EOL_RAW %>% 
  ggplot() +
  aes(x=datetime, y=salinity) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Salinity") +
  ggtitle("EOL time series observations of salinity  2013-2022")

plot_salinity
plot_ano_sal
```


## Part 2 : Trend analysis by autocorrelation



